---
title: "Assignment 1: Temperature and DO"
author: "Sam Struthers & Carolina Barbosa - adapted from Nick Gubbins"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dataRetrieval)
library(lubridate)
library(tidyverse)
```

# Ecological Context

## Temperature as a master control

Available energy in a system is the master control on which, and at what
rate, chemical reactions take place. Temperature is the most common, and
most easily accessible, measurement of energy in a system. While other
forms of energy ( such as turbulence) are critical for understanding
aquatic systems, temperature can be easily and cheaply measured without
effecting the system itself. For this week's assignment, we will be
getting more familiar with R by downloading and analyzing temperature
data from the USGS.

## Power generation and water temperature

Rivers are often used as a water source for the production of steam in
power generation. Generating stations draw in water from upstream, use
it to create steam, use that steam to spin turbines, then (mostly) cool
the steam, and re-release it downstream. You can clearly see the effect
of this on the White River in downtown Indianapolis, IN. Station
03353000 is located just upstream of a major generating plant. Station
03353611 is just below it. The gauges are only \~2 miles apart. Let's
load in a few months of data from both and see how they differ.

```{r}

site_data <- readNWISuv(siteNumbers = c( '03353000', "03353611"),
                   parameterCd = '00010',
                   startDate = '2024-10-01',
                   endDate =  '2025-01-01')

ggplot(site_data, aes(x = dateTime, y = X_00010_00000, color = site_no))+
  geom_line()+
  labs(x = 'Date', y = 'Temperature (degrees C)', color = 'Site')

```

You can see the effect of the generating station most clearly as
temperatures drop for the winter. Water is usually returned at \~20
degrees C, giving the downstream site a higher average temperature.

## Dissolved oxygen and temperature

Dissolved oxygen is a key indicator of aquatic health. It is produced by
photosynthesizing microbes and macrophytes. Heterotrophs, like fish and
many kinds of bacteria, use dissolved oxygen in respiration. Healthy
freshwater systems tend to have dissolved oxygen levels above 6 mg/L
(usually around 80% saturation or higher). When dissolved oxygen levels
get too low (below 6 mg/L) fish cannot spawn, as fish eggs rely on
passive diffusion of dissolved oxygen to survive. Below 5 mg/L DO, you
can see dramatic [fish
kills](https://apnews.com/article/837ec91d3a1c4f86b1258be93cd865e8).
These kills represent a huge disruption to the river's food chain and a
massive influx of organic matter, often leading to spiraling water
quality issues.

# Assignment

Read in **mean daily** discharge (code '00060'), DO (code '00300'), and
temperature (code '00010') data from the South Platte River at
Englewood, CO ('06711565') for the 2023 water year

Let's first look at the readNWISdv function to see what it does.

```{r}
#?readNWISdv
```

Now we can will in the function inputs to get the data we want

```{r}
data <- readNWISdv(siteNumbers = '06711565',
                   parameterCd = c('00060', '00300', '00010'),
                   startDate = '2023-10-01',
                   endDate =  '2024-9-30')
```

## Q1 (10pts):

Look at the structure of the data in any way you choose.

## Q2 (10 pts):

Rename your columns to something sensible, omit rows without complete
observations (NAs), and reduce your data frame to only include the date
and your data of interest.

## Q3 (10pts):

Plot a time series of all three parameters in the same graph.

## Q4 (10pts):

This time series graph is not very useful to us. Why not? Give two
examples of design choices you could make to improve its usefulness.

## Q5 (20pts):

Implement both of your proposed changes.

## Q6 (10pts):

Scatterplots are a great way to compare correlations. Create two
scatterplots. One of DO vs Q, the other of DO vs temperature.

## Q7 (10pts):

Use the 'cor()' function to test which two variables are the most
closely related (look at the 'help' window for documentation). Which
relationship is it? Why do you think that is? Report your answers in
complete sentences.

## Q8 (30 pts):

Imagine the city of Englewood is exploring building a coal-fired power
plant using the South Platte as its water source and you have been hired
on as a consultant. **Write a paragraph** brief for the city. Given that
the benefits to the city are concrete, what issues would you raise? What
effects do you think the plant would have on water temperatures and
water quality? What data/assurances would you need to see before
thinking the project should go ahead?

## Bonus Q ([**Ungraded**]{.underline}):

Pull in daily discharge (code '00060') data for the two USGS sites on
the Cache la Poudre River in Fort Collins, CO ('06752260' and
'06752280') for the 2023 water year (2023-10-01 - 2024-09-30). Plot the
data in a way that shows the difference between the two sites. What do
you think is causing the difference?
