---
title: "Tidy Tuesday"
author: "Sam Struthers"
date: "`r Sys.Date()`"
output: html_document
---

# Set up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
##install.packages("tidytuesdayR")

today <- Sys.Date()+ days(1)

tuesdata <- tidytuesdayR::tt_load(today)

names(tuesdata)

```

# Example for 2/4/24

## Make a plot of IMDB ratings over time

```{r}
episodes <- tuesdata$simpsons_episodes

ggplot(episodes, aes(x = original_air_date, y = imdb_rating)) +
  geom_point() +
  #geom_smooth() +
  labs(title = "IMDB Ratings of Simpsons Episodes Over Time",
       x = "Air Date",
       y = "IMDB Rating (0-10)")

```

## Do more people vote for better episodes?

Lets see if there is a correlation with the number of votes and IMDB scores

```{r}

ggplot(episodes, aes(x = imdb_votes, y = imdb_rating)) +
  geom_point() +
  #geom_smooth() +
  labs(title = "IMDB Ratings of Simpsons Episodes with number of votes",
       x = "Number of Votes",
       y = "IMDB Rating (0-10)")

```

## Are the most watched shows the highest rated?

We would think that the most people would watch the best episodes, but is that true?

```{r}
ggplot(episodes, aes(x = us_viewers_in_millions , y =  imdb_rating)) +
  geom_point() +
  #geom_smooth() +
  labs(title = "IMDB Ratings of Simpsons Episodes with number of viewers",
       x = "Number of US viewers (Millions)",
       y = "IMDB Rating (0-10)")

summary(lm(episodes$imdb_rating ~ episodes$us_viewers_in_millions))


```

## Maybe the most watched episodes are the season premiere or season finale?

```{r}

episodes <- episodes%>%
  group_by(season) %>% #group by the season
  mutate(order = ifelse(number_in_season == 1, "premiere", # the premiere is always the first
                        ifelse(max(number_in_season) == number_in_season, "finale", # the finale is the last of the season
                                "normal")))%>% # all other episodes
  ungroup()
#lets rearrange these so that our plot looks nice
episodes$order = factor(episodes$order, levels = c("premiere",  "normal", "finale"))

ggplot(episodes, aes(x = order,  y = us_viewers_in_millions, color = order)) +
  geom_boxplot() +
  labs(title = "Number of US viewers for Simpsons Episodes by Episode Type",
       x = "Episode Type",
       y = "Number of US viewers (Millions)", 
       color = "")+
  theme_bw()

# test the difference between the finale and premiere
t.test(us_viewers_in_millions ~ order, data = episodes%>%filter(order %in% c("premiere", "finale")))


```




