---
title: "R Tips"
author: "Sam Struthers"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

# Set up

We will typically packages to accomplish alot of the tasks in this
class. To use the package, you need to first install it (only once) and
then library it (everytime you start a new R session).

```{r setup, message = FALSE}
#install.packages("dataRetrieval")
#install.packages("lubridate")
#install.packages("tidyverse")

library(dataRetrieval)
library(lubridate)
library(tidyverse)
```

# Packages

Tidyverse: Tidyverse is an collection of R packages that work together
to make data transformation and handling easier. Please use them!
Lubridate: Lubridate is a package that makes working with dates and
times easier. It is part of the Tidyverse. GGplot2: GGplot2 is a package
that makes it easier to create plots in R. It is part of the Tidyverse.
Arrow: Arrow is a package to quickly read in various forms of data
DataRetrieval: DataRetrieval is a package that makes it easier to
retrieve data from the USGS NWIS database.

# Quick Tips

To create a new chunk (like the one below), you can use the keyboard
shortcut Ctrl + Alt + I (Cmd + Option + I on a Mac).

```{r}
# This is a code chunk
```

To run a line of code, you can use the keyboard shortcut Ctrl + Enter
(Cmd + Enter on a Mac).

```{r}
#To comment out code, use the # symbol in code chunks.
#This will prevent the code from running.

#To comment out multiple lines of code, highlight the lines and use the 
#keyboard shortcut Ctrl + Shift + C (Cmd + Shift + C on a Mac).
print("I love commenting my code!!") # You can put comments after a code like this

print("I love commenting my code!!") # Ex: This is how I can print text in R

```

Parentheses/curly brackets can get a bit confusing in long functions.
Engage rainbow parentheses by going to Tools -\> Global Options -\> Code
-\> Display -\> Show rainbow parentheses.

```{r}
for (i in 1:10) {
  if (i %% 2 == 0) {
    print(i)
  }
}
```

You can also change the color of your R Console by going to Tools -\>
Global Options -\> Appearance -\> Editor theme. (Personally I like
Merbivore Soft)

# Reproducible Examples

Helpful Example
- Shows packages used
- Shows/loads the dataset
- Then shows the code you are trying to troubleshoot

```{r}
library(ggplot2)
library(dplyr)

# Show/Load dataset
df <- tibble::tibble(
  category = c("A", "B", "C", "D"),
  value = c(10, 15, 7, 20)
)

# Attempt to plot the data
ggplot(df, aes(x = category, y = value)) +
  geom_col(fill = "blue") +
  labs(title = "Category vs Value")

```

Not Helpful Example

```{r}
ggplot(my_data, aes(x = category, y = value)) +
  geom_col()

```

